Partridge: An effective system for the automatic classification of the types of academic papers

### 2013-07-16 doc/bcspaper/article.tex
Fixed 6 pages issue with vspaces, horrible but it worked

\vspace{-15pt}
\section{Introduction} \label{sec:1}

Since the advent of the `Digital Age', the amount of information available to
researchers has been increasing drastically, relevant material is becoming
progressively more difficult to find manually and the need for an automated
information retrieval tool more apparent. There are already a large number of
information retrieval and recommendation systems for scientific publications.
Many of these systems, such as AGRICOLA \footnote{\url{http://agricola.nal.usda.gov/}},
the Cochrane Library\footnote{\url{http://www.thecochranelibrary.com/}} and Textpresso
\footnote{\url{http://www.textpresso.org/}} index only publications from predefined
journals or topics (for the above examples, Agriculture, Biology and
Bioinformatics respectively).  Unfortunately, these domain specific indexing
systems usually only contain a small subset of papers, excluding potentially
crucial literature because it does not quite fit into the subject domain.

@@

\subsection*{Collection of scientific articles} Partridge allows users to
upload any paper which is free of copyright restrictions.  For the purpose of
this study we needed a large set of papers that we could label with CoreSC to
investigate how this information assists classification into paper type.  Open
Access (OA) journals provide free read access to their papers, but many do not
permit the user of articles for data mining purposes. The Public Library of
Science (PLoS) journals contain large volumes of OA literature under a
permissive license that allows data mining.  They also use the PubMed Central
markup schema, which is compatible with SAPIENTA, for papers published through
their journal. The PLoS journals advanced search offers approximately 50 types
of paper through which to restrict the search.  Many of these paper type
categories contain too few papers to be useful, others are too ad hoc (e.g.
{\it Message from PLoS}). Indeed it is not clear how these paper types have
been identified. We chose to look at a range of types, some of which we expect
to overlap or have an unclear distinction.  These are namely: Essay,
Correspondence, Case Study, Perspective, Viewpoint, Opinion, Review, Research.

@@

\subsection*{Paper Processing and classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting) 
In order to obtain automatic labeling of sentences in a paper with
CoreSC concepts we first needed to convert the papers to a format that
the the CoreSC classifier, SAPIENTA can analyse.  Currently, SAPIENTA
supports the SciXML and PubMed Central DTDs.  Papers in PDF format are
first converted to XML using PDFX %\cite{PDFX} - no paper known yet
, a free service hosted by the University of Manchester
(\cite{Constantin2013}). Once the paper has been split into
sentences, SAPIENTA is used to annotate each sentence with a relevant
CoreSC label. These labels are assigned using a conditional random
fields model based upon a number of features such as the sentence's
location within the paper and pairs and triplets of words found
consecutively within the sentence \cite{Liakata2012}.  The SAPIENTA
system has been trained on a corpus of 265 chemistry and biochemistry
papers and we have done no domain adaptation for the papers in PLoS.

\begin{figure}[t]

\centering

%\subfloat{
\includegraphics[width=0.75\textwidth]{figures/corescs.png}
%}
%\subfloat{
%\includegraphics[width=0.45\textwidth]{figures/report_corescs.png}
%}
\caption{The CoreSC content of a research paper (left) and a review paper (right) randomly selected from the corpus}

\vspace{-20pt}
\label{fig:coresc_pies} 
\end{figure}

@@

\vspace{-15pt}
\section{Results and Discussion}
\label{sec:3}

@@

\caption{Per-class recall, precision and F-measure micro-averaged over 10-fold cross validation, reporting the results on the held-out validation segment. Top paper types in bold.}

@@

We inspected the detail of a single decision tree, grown on the entire dataset,
to give further insight into which CoreSC class decisions were responsible for
the paper type predictions. This was a very large tree, with a depth of 37
nodes in places.  The first decision was regarding the number of Background
sentences in the article. A low Background percentage, of less than 0.694
indicates a Correspondence paper. 28 out of the 31 papers that were classified
by this decision as Correspondence were correctly labelled.

@@

\vspace{-15pt}
\section{Summary and Conclusions}
\label{sec:4}

To summarise, we have demonstrated that paper type can largely be predicted
from the distribution of the automatically obtained CoreSC sentence labels in
the full text of a paper. We have described some of the particular CoreSC
features that determine a paper type (e.g. Correspondence characterised by very
little Background, Opinions characterised by low percentage of Experiment) and
discussed which of the paper types are not easily separable in this way. We
recommended for example merging the Opinion, Viewpoint and Perspective articles
into a single category, which increased overall classification performance. It
is also potentially useful to distinguish between two types of Case study.
Analysis of an example tree shows the decision making process to be complex,
but agrees with our general understanding of paper types. The Partridge system
allows refinement of paper search using CoreSC and paper type, both of which
can be intelligently determined using machine learning methods.

@@

\begin{acknowledgement} We thank the Leverhulme trust for the support to Dr
Liakata's Early Career Fellowship and also EMBL-EBI, Cambridge UK for the
facilities offered to Dr Liakata.  \end{acknowledgement}

\vspace{-15pt}
\bibliographystyle{plain}
\bibliography{partridge}

>>>>

\vspace{-15pt}
\section{Introduction} \label{sec:1}
\vspace{-11pt}
Since the advent of the `Digital Age', the amount of information available to
researchers has been increasing drastically, relevant material is becoming
progressively more difficult to find manually and the need for an automated
information retrieval tool more apparent. There are already a large number of
information retrieval and recommendation systems for scientific publications.
Many of these systems, such as AGRICOLA\footnote{\url{http://agricola.nal.usda.gov/}},
the Cochrane Library\footnote{\url{http://www.thecochranelibrary.com/}} and Textpresso
\footnote{\url{http://www.textpresso.org/}} index only publications from predefined
journals or topics (for the above examples, Agriculture, Biology and
Bioinformatics respectively).  Unfortunately, these domain specific indexing
systems usually only contain a small subset of papers, excluding potentially
crucial literature because it does not quite fit into the subject domain.

@@

\subsection*{Collection of scientific articles} 
\vspace{-15pt}
Partridge allows users to upload any paper which is free of copyright
restrictions.  For the purpose of this study we needed a large set of papers
that we could label with CoreSC to investigate how this information assists
classification into paper type.  Open Access (OA) journals provide free read
access to their papers, but many do not permit the user of articles for data
mining purposes. The Public Library of Science (PLoS) journals contain large
volumes of OA literature under a permissive license that allows data mining.
They also use the PubMed Central markup schema, which is compatible with
SAPIENTA, for papers published through their journal. The PLoS journals
advanced search offers approximately 50 types of paper through which to
restrict the search.  Many of these paper type categories contain too few
papers to be useful, others are too ad hoc (e.g.  {\it Message from PLoS}).
Indeed it is not clear how these paper types have been identified. We chose to
look at a range of types, some of which we expect to overlap or have an unclear
distinction.  These are namely: Essay, Correspondence, Case Study, Perspective,
Viewpoint, Opinion, Review, Research.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.75\textwidth]{figures/corescs.png}
\caption{The CoreSC content of a research paper (left) and a review paper (right) randomly selected from the corpus}
\vspace{-20pt}
\label{fig:coresc_pies} 
\end{figure}

\subsection*{Paper Processing and classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting) 
\vspace{-15pt}

In order to obtain automatic labeling of sentences in a paper with
CoreSC concepts we first needed to convert the papers to a format that
the the CoreSC classifier, SAPIENTA can analyse.  Currently, SAPIENTA
supports the SciXML and PubMed Central DTDs.  Papers in PDF format are
first converted to XML using PDFX %\cite{PDFX} - no paper known yet
, a free service hosted by the University of Manchester
(\cite{Constantin2013}). Once the paper has been split into
sentences, SAPIENTA is used to annotate each sentence with a relevant
CoreSC label. These labels are assigned using a conditional random
fields model based upon a number of features such as the sentence's
location within the paper and pairs and triplets of words found
consecutively within the sentence \cite{Liakata2012}.  The SAPIENTA
system has been trained on a corpus of 265 chemistry and biochemistry
papers and we have done no domain adaptation for the papers in PLoS.

@@

\vspace{-15pt}
\section{Results and Discussion}
\label{sec:3}
\vspace{-15pt}

@@

\vspace{-10pt}
\caption{Per-class recall, precision and F-measure micro-averaged over 10-fold cross validation, reporting the results on the held-out validation segment. Top paper types in bold.}

@@

We inspected the detail of a single decision tree, grown on the entire dataset,
to give further insight into which CoreSC class decisions were responsible for
the paper type predictions. This was a very large tree, with a depth of 37
nodes in places.  The first decision was based upon the number of Background
sentences in the article. Low Background percentages, of less than 0.694
indicates a Correspondence paper. 28 out of the 31 Correspondence papers were
correctly labelled by this decision  classified by this decision as
Correspondence were correctly labelled.

@@

\vspace{-15pt}
\section{Summary and Conclusions}
\label{sec:4}
\vspace{-15pt}
To summarise, we have demonstrated that paper type can largely be predicted
from the distribution of the automatically obtained CoreSC sentence labels in
the full text of a paper. We have described some of the particular CoreSC
features that determine a paper type (e.g. Correspondence characterised by very
little Background, Opinions characterised by low percentage of Experiment) and
discussed which of the paper types are not easily separable in this way. We
recommended for example merging the Opinion, Viewpoint and Perspective articles
into a single category, which increased overall classification performance. It
is also potentially useful to distinguish between two types of Case study.
Analysis of an example tree shows the decision making process to be complex,
but agrees with our general understanding of paper types. Partridge allows refinement of paper search using CoreSC and paper type, both of which
can be intelligently determined using machine learning methods.

@@

\vspace{-10pt}
\begin{acknowledgement} We thank the Leverhulme trust for the support to Dr
Liakata's Early Career Fellowship and also EMBL-EBI, Cambridge UK for the
facilities offered to Dr Liakata.  \end{acknowledgement}
\vspace{-2pt}

\vspace{-30pt}
\bibliographystyle{plain}
\bibliography{partridge}
